---
title: "Bayesian mediation models for number of species and number of canopy species"
date: "20/12/2022"
output: word_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(here)
library(future)
library(brms)

options(digits=3)

```

# Canopy cover mediates the effects of a decadal increase in time since fire on arboreal birds

Complete and partial Bayesian mediation models for the number of species and the number of canopy species.     

*Study area*  
The Greater Blue Mountains World Heritage Area, Australia.  
46 33-ha circular sites.  

*Site level*  
TotalSp: the total number of species detected in each site (analytical code provided below).    
CanForSp; the number of canopy-foraging species in each site (same code as above, not shown).    

TSF; an indicator variable for time since fire, 0 if five years (last burnt 2013), 1 if 16 years (last burnt 2002).         
CPY; the proportion of canopy cover over each site.    
LAT; latitude at the center of each site (standardised, mean = 0, s.d. = 1).  
SEV; the proportion of each site that had been burnt by high-severity fire (standardised).    


### Read in the data

```{r}

path <- here("data", "dat.mod.csv")
dat.mod <- read.csv(path, stringsAsFactors = FALSE)

```


### Count model: complete mediation

This model represents the TSF effect on the number of species as being completely mediated through canopy cover.  
Equation 1 in the article.

```{r}

canopy_mod <- bf(CPY ~ LAT + SEV + TSF, 
                   family = Beta(link = "logit"))

count_mod_complete <- bf(TotalSp ~ LAT + SEV + CPY, 
                   family = brmsfamily("com_poisson", link = "log", link_shape = "log"))

prior <- set_prior("normal(0,10)", class = "b", resp = c("TotalSp", "CPY")) +
                   set_prior("normal(0,10)", class = "Intercept", resp = c("TotalSp", "CPY"))
          

future::plan(multisession, workers = 4)
  
 count_complete_result <-  brm(
    count_mod_complete + canopy_mod + set_rescor(FALSE),
    prior = prior,
    save_pars = save_pars(all = TRUE),  # to enable moment_match for LOO
    data = dat.mod,
    chains = 4, 
    iter = 11000, 
    warmup = 1000, 
    thin = 10, 
    future = TRUE, 
    control = list(adapt_delta = 0.95),
    inits = "0"
  )
  
future::plan(sequential)

save(count_complete_result, file = 'models/count_complete_result.RData')

load('models/count_complete_result.RData')

# Check of the fitted model terms and number of effective samples
summary(count_complete_result)
mcmc_plot(count_complete_result, type = "dens")

# Obtain Bayesian R2 values
bayes_R2(count_complete_result, probs = c(0.05, 0.95))

```


### Count model: partial mediation

This model represents the TSF effect on the number of species as being partially mediated through canopy cover.  
Equation 2 in the article.  

```{r}

canopy_mod <- bf(CPY ~ LAT + SEV + TSF, 
                   family = Beta(link = "logit"))

count_mod_partial  <- bf(TotalSp ~ LAT + SEV + CPY + TSF, 
                   family = brmsfamily("com_poisson", link = "log", link_shape = "log"))

prior <- set_prior("normal(0,10)", class = "b", resp = c("TotalSp", "CPY")) +
                   set_prior("normal(0,10)", class = "Intercept", resp = c("TotalSp", "CPY"))

future::plan(multisession, workers = 4)

  count_partial_result <-  brm(
    count_mod_partial + canopy_mod + set_rescor(FALSE),
    prior = prior,
    save_pars = save_pars(all = TRUE),  # to enable moment_match for LOO
    data = dat.mod,
    chains = 4, 
    iter = 11000, 
    warmup = 1000, 
    thin = 10, 
    future = TRUE, 
    control = list(adapt_delta = 0.95),
    inits = "0"
  )
  
future::plan(sequential)
  
save(count_partial_result, file = 'models/count_partial_result.RData')
  
load('models/count_partial_result.RData')
  
# Check of the fitted model terms and number of effective samples
summary(count_partial_result)
mcmc_plot(count_partial_result, type = "dens")

# Obtain Bayesian R2 values
bayes_R2(count_partial_result, probs = c(0.05, 0.95))  

```


### Model comparison
Compare complete and partial mediation models using LOO via brms

```{r}

x <- LOO(count_complete_result, count_partial_result, moment_match = TRUE) # moment matching applied as per methods  

# Difference in ELPD values between models
x$diffs

```
