---
title: "Bayesian mediation models for the occurrence of individual canopy species"
date: "20/12/2022"
output: word_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(here)
library(future)
library(brms)

options(digits=3)

```

# Canopy cover mediates the effects of a decadal increase in time since fire on arboreal birds

Complete and partial Bayesian mediation models for the occurrence of individual canopy species.         

*Study area*  
The Greater Blue Mountains World Heritage Area, Australia.    
46 33-ha circular sites.    

*Site level*  
The following canopy species were detected/not detected (1/0) in each site.  
Analytical code was the same for each species, with code for the BHHE provided below as an example.  
BHHE;	Melithreptus brevirostris,	Brown-headed Honeyeater.    
CBCK;	Scythrops novaehollandiae,	Channel-billed Cuckoo.    
CICB;	Coracina tenuirostris,	Cicadabird.    
GGCT;	Callocephalon fimbriatum,	Gang-Gang Cockatoo.    
LEFC;	Myiagra rubecula,	Leaden Flycatcher.    
MTBD;	Dicaeum hirundinaceum,	Mistletoebird.    
OBOR;	Oriolus sagittatus,	Olive-backed Oriole.    
STTB;	Acanthiza lineata,	Striated Thornbill.    

TSF; an indicator variable for time since fire, 0 if five years (last burnt 2013), 1 if 16 years (last burnt 2002).           
CPY; the proportion of canopy cover over each site.      
LAT; latitude at the center of each site (standardised, mean = 0, s.d. = 1).    
SEV; the proportion of each site that had been burnt by high-severity fire (standardised).      


### Read in the data

```{r}

path <- here("data", "dat.mod.csv")
dat.mod <- read.csv(path, stringsAsFactors = FALSE)

```


### Occurrence model: complete mediation

This model represents the TSF effect on the occurrence of individual canopy species as being completely mediated through canopy cover.      
Equation 3 in the article.  

```{r}

canopy_mod <- bf(CPY ~ LAT + SEV + TSF, 
                   family = Beta(link = "logit"))

occur_mod_complete <- bf(BHHE ~ LAT + SEV + CPY, 
                                 family = bernoulli(link = "logit"))

prior <- set_prior("normal(0,10)", class = "b", resp = c("BHHE", "CPY")) +
                   set_prior("normal(0,10)", class = "Intercept", resp = c("BHHE", "CPY"))

future::plan(multisession, workers = 4)
  
 occur_complete_result <-  brm(
    occur_mod_complete + canopy_mod + set_rescor(FALSE),
    prior = prior,
    save_pars = save_pars(all = TRUE),  # to enable moment_match for LOO
    data = dat.mod,
    chains = 4, 
    iter = 11000, 
    warmup = 1000, 
    thin = 10, 
    future = TRUE, 
    control = list(adapt_delta = 0.95),
    inits = "0"
  )
  
future::plan(sequential)

save(occur_complete_result, file = 'models/occur_complete_result.RData')

load('models/occur_complete_result.RData')

# Check of the fitted model terms and number of effective samples
summary(occur_complete_result)
mcmc_plot(occur_complete_result, type = "dens")

# Obtain Bayesian R2 values
bayes_R2(occur_complete_result, probs = c(0.05, 0.95))

```


### Occurrence model: partial mediation

This model represents the TSF effect on the occurrence of individual canopy species as being partially mediated through canopy cover.      
Equation 4 in the article.    

```{r}

canopy_mod <- bf(CPY ~ LAT + SEV + TSF, 
                   family = Beta(link = "logit"))

occur_mod_partial  <- bf(BHHE ~ LAT + SEV + CPY + TSF, 
                      family = bernoulli(link = "logit"))

prior <- set_prior("normal(0,10)", class = "b", resp = c("BHHE", "CPY")) +
                   set_prior("normal(0,10)", class = "Intercept", resp = c("BHHE", "CPY"))

  future::plan(multisession, workers = 4)

  occur_partial_result <-  brm(
    occur_mod_partial + canopy_mod + set_rescor(FALSE),
    prior = prior,
    save_pars = save_pars(all = TRUE),  # to enable moment_match for LOO
    data = dat.mod,
    chains = 4, 
    iter = 11000, 
    warmup = 1000, 
    thin = 10, 
    future = TRUE, 
    control = list(adapt_delta = 0.95),
    inits = "0"
  )
  
future::plan(sequential)

save(occur_partial_result, file = 'models/occur_partial_result.RData')
  
load('models/occur_partial_result.RData')

# Check of the fitted model terms and number of effective samples
summary(occur_partial_result)
mcmc_plot(occur_partial_result, type = "dens")

# Obtain Bayesian R2 values
bayes_R2(occur_partial_result, probs = c(0.05, 0.95))

```


### Model comparison
Compare complete and partial mediation models using LOO via brms

```{r}

x <- LOO(occur_complete_result, occur_partial_result, moment_match = TRUE) # moment matching applied as per methods  

# Difference in ELPD values between models
x$diffs

```

